### 2016-11-22 Paper_v9.tex
start of reformmating revisions

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2009}

@@

@@

%as seen in the training objectives.

@@

\subsection*{TGF$-\beta$1/2 and VEGF-A exposure promotes phenotype heterogeneity through NFATc and phosphorylated Sp1}
While we captured the central tendency of many of the molecular features of EMT induction following TGF$-\beta$1/2 exposure, an often neglected but important emergent
feature of developmental and pathological programs is population heterogeneity \cite{Park:2010fu}.
We (and others) have previously hypothesized that deterministic model ensembles can simulate population behavior, at least at a course grained level \cite{Lequieu:2011fj}.
We tested this hypothesis by analyzing the response of the population of EMT models to extracellular cues and then comparing this response to flow cytometry studies.
We quantified the phenotypic response of the individual members of the ensemble to TGF$-\beta$1/2 stimulation for two downstream phenotypic markers, Vimentin (mesenchymal) and E-cadherin (epithelial) following the addition of TGF$-\beta$1/2 alone (Fig. \ref{fg:F3}), and/or VEGF-A in combination with NFATc inhibitors (Fig. \ref{fg:F3}).
%We were further able to use this analysis to examine the activity and effects of underlying signaling molecules.

We identified model subpopulations that exhibited different behaviors following exposure to TGF$-\beta$1/2 (Fig. \ref{fg:F3}B).
Analysis of the molecular signatures of these subpopulations suggested the abundance, localization and state of the Sp1, AP-1 and NFATc transcription factors controlled population heterogeneity.
The majority of models ($>$80\%) responded to treatment, moving away from the untreated population (Fig \ref{fg:F3}A-F, gray).
These models showed the classically expected behavior, a switch from an epithelial to mesenchymal phenotype following TGF$-\beta$1/2 exposure.
Some models resembled untreated cells; they had elevated phosphorylated Sp1, relative to non-induced cells,
which decreased E-cadherin expression through Slug-mediated inhibition, which in turn increased Vimentin expression through TGF$-\beta$3 autocrine signaling and the liberation of $\beta$-catenin. However, the most biologically interesting behavior was exhibited by cells achieving a hybrid phenotype, most notable in a dual treatment condition (\ref{fg:F3}C, black arrow), but also present in a small percentage of untreated cells (Fig. \ref{fg:F3}B, gray arrow).
Models with this hybrid phenotype had elevated Sp1 and NFAT transcriptional activity, resulting in simultaneously increased Vimentin and E-cadherin expression (Fig. \ref{fg:F4}A).
However, these conclusions are likely sensitivity to the components that we have included in the model, and could also involve proteins that we have not considered.

%Analysis of these hypothetical cells suggested they had deregulated NFAT expression and nuclear localization promoted E-cadherin expression while TGF$-\beta$1/2 induced Sp1 action promoted Vimentin expression.

%In particular, when present, VEGF-A signaling through NFATc was predicted to provide a protective up-regulation of E-cadherin.

@@

@@

% \begin{equation}\label{asseassertain_massbalance}
% \frac{d\mathbf{x}}{dt}=\mathbf{S}\cdot\mathbf{r}\left(\mathbf{x},\mathbf{k}\right)\qquad\mathbf{x}\left(t_{o}\right)=\mathbf{x}_{o}
% \end{equation}
% The quantity $\mathbf{x}$ denotes the vector describing the abundance of protein, mRNA, and other species in the model ($995\times 1$).
% The stoichiometric matrix $\mathbf{S}$ encodes the signaling architecture considered in the model ($995\times 1700$).
% Each row of $\mathbf{S}$ describes a signaling component while each column describes a particular interaction.
% The $(i,j)$ element of $\mathbf{S}$, denoted by $\sigma_{ij}$, describes how species $i$ is involved with interaction $j$.
% If $\sigma_{ij}>0$, species $i$ is produced by interaction $j$. Conversely, If $\sigma_{ij}<0$, then species $i$ is consumed in interaction $j$.
% Lastly, if $\sigma_{ij}=0$, then species $i$ is not involved in interaction $j$.
% The term  $\mathbf{r}\left(\mathbf{x},\mathbf{k}\right)$ denotes the vector of interactions rates ($1700\times{1}$).
% We modeled each network interaction (gene expression, translation and biochemical transformations)
% using elementary rate laws where all reversible interactions were split into two irreversible steps (supplemental materials).
% Thus, the rate expression for interaction $q$ was given by:
% \begin{equation}\label{eq-mass-action}
% r_{q}\left(\mathbf{x},k_{q}\right)=k_{q}\prod_{j\in\left\{\mathbf{R}_{q}\right\}}x_{j}^{-\sigma_{jq}}
% \end{equation}
% The set $\left\{\mathbf{R}_{q}\right\}$ denotes reactants for reaction $q$, while $\sigma_{jq}$ denotes the stoichiometric coefficient (element of the matrix $\mathbf{S}$) governing species $j$ in reaction $q$. The quantity $k_{q}$ denotes the rate constant (unknown) governing reaction $q$.
% Model equations were generated in the C-programming language using the UNIVERSAL code generator, starting from an text-based input file (supplemental materials).
% UNIVERSAL, an open source Objective-C/Java code generator, is available as a Google Code project (http://code.google.com/p/universal-code-generator/).
% Model equations were solved using the CVODE solver in the SUNDIALS library \cite{Hindmarsh2005} on an Apple workstation (Apple, Cupertino, CA)
% as previously described \cite{Tasseff:2011ve}.

@@

% \subsubsection*{Robustness coefficients.} Robustness coefficients were calculated as shown previously \cite{Tasseff:2011ve,Lequieu:2011fj}.
% Robustness coefficients denoted by $\alpha\left(i,j,t_{o},t_{f}\right)$:
% \begin{equation}
% \alpha\left(i,j,t_{o},t_{f}\right)=\left(\displaystyle\int_{t_{o}}^{t_{f}}x_{i}\left(t\right)dt\right)^{-1}\left(\displaystyle\int_{t_{o}}^{t_{f}}x^{(j)}_{i}\left(t\right)dt\right)
% \end{equation}
% quantify the response of a marker to a structural or operational perturbation to the network architecture.
% Here $t_o$ and $t_f$ denote the initial and final simulation time respectively, while
% $i$ and $j$ denote the indices for the marker and the perturbation respectively. A value of $\alpha\left(i,j,t_{o},t_{f}\right)>1$, indicates increased marker abundance, while $\alpha\left(i,j,t_{o},t_{f}\right)<1$ indicates decreased marker abundance following perturbation $j$. If
% $\alpha\left(i,j,t_{o},t_{f}\right)\sim{1}$ the $j$th perturbation does not influence the abundance of marker $i$.
% Robustness coefficients were calculated for each member of the ensemble (N $\simeq$ 1100).

@@

%\section*{Acknowledgments}
%The project described was supported by Award Number \#U54CA143876 from the National Cancer Institute to JDV, National Science Foundation CBET-0955172 to JTB, and National Institute of Health HL110328 to JTB. The content is solely the responsibility of the authors and does not necessarily represent the official views of the National Cancer Institute or the National Institutes of Health. Lastly, we thank the anonymous reviewers for their helpful suggestions.

\clearpage
% \bibliographystyle{plain}
%\bibliographystyle{IEEEbib}

>>>>

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

@@

\subsection*{TGF$-\beta$1/2 and VEGF-A exposure promotes phenotype heterogeneity through NFATc and phosphorylated Sp1}
While we captured the central tendency of many of the molecular features of EMT induction following TGF$-\beta$1/2 exposure, an often neglected but important emergent
feature of developmental and pathological programs is population heterogeneity \cite{Park:2010fu}.
We (and others) have previously hypothesized that deterministic model ensembles can simulate population behavior, at least at a course grained level \cite{Lequieu:2011fj}.
We tested this hypothesis by analyzing the response of the population of EMT models to extracellular cues and then comparing this response to flow cytometry studies.
We quantified the phenotypic response of the individual members of the ensemble to TGF$-\beta$1/2 stimulation for two downstream phenotypic markers, Vimentin (mesenchymal) and E-cadherin (epithelial) following the addition of TGF$-\beta$1/2 alone (Fig. \ref{fg:F3}), and/or VEGF-A in combination with NFATc inhibitors (Fig. \ref{fg:F3}).

We identified model subpopulations that exhibited different behaviors following exposure to TGF$-\beta$1/2 (Fig. \ref{fg:F3}B).
Analysis of the molecular signatures of these subpopulations suggested the abundance, localization and state of the Sp1, AP-1 and NFATc transcription factors controlled population heterogeneity.
The majority of models ($>$80\%) responded to treatment, moving away from the untreated population (Fig \ref{fg:F3}A-F, gray).
These models showed the classically expected behavior, a switch from an epithelial to mesenchymal phenotype following TGF$-\beta$1/2 exposure.
Some models resembled untreated cells; they had elevated phosphorylated Sp1, relative to non-induced cells,
which decreased E-cadherin expression through Slug-mediated inhibition, which in turn increased Vimentin expression through TGF$-\beta$3 autocrine signaling and the liberation of $\beta$-catenin. However, the most biologically interesting behavior was exhibited by cells achieving a hybrid phenotype, most notable in a dual treatment condition (\ref{fg:F3}C, black arrow), but also present in a small percentage of untreated cells (Fig. \ref{fg:F3}B, gray arrow).
Models with this hybrid phenotype had elevated Sp1 and NFAT transcriptional activity, resulting in simultaneously increased Vimentin and E-cadherin expression (Fig. \ref{fg:F4}A).
However, these conclusions are likely sensitive to the components that we have included in the model, and could also involve proteins that we have not considered.

@@

@@

@@

%\clearpage

@@

@@

@@
